---
title: "Poisson(lambda=x) vs Normal(mean=x, sd=sqrt(x))"
author: "Mauricio Alarcon"
date: "March 12, 2015"
output: html_document
---

For any given dataset with mean=x and sd=sqrt(x), both the Poisson and Normal distributions can be good pdfs.

In this example, I'm comparing poisson and normal distributions with the following paramenters:

* **Mean:** 25
* **Standard Deviation:** 5
* **Lambda:** 25

Both Poisson(lambda=25) and Normal(mean=25, sd=5) are awfully similar as we can see below:

* * * 
### Summary
```{r, kable,echo=FALSE, warning=FALSE, message=FALSE}
library(ggplot2)
library(reshape)
library(knitr)

range<-c(5:45)
mean<-25
sd<-sqrt(mean)
#load poisson data
distributions <- data.frame(dist=sprintf("Poisson(lambda=%i)",mean)
                            ,x=range
                            ,y=dpois(range,mean)
                            )
# add normal data
distributions <- rbind(distributions
                       ,data.frame(dist=sprintf("Normal(mean=%i, sd=%i)",mean,sd)
                                   ,x=range
                                   ,y=dnorm(range,mean,sd)
                                   )
                       )


options(digits=4)
kable(summary(cast(distributions, x ~ dist)))
```

* * * 

###Histograms
```{r, echo=FALSE}
qplot(x,y,data=distributions
      ,color=dist
      ,facets=dist~.
      ,fill=dist
      ,geom=c( "line")) + theme(legend.position="none")
```

* * * 

### Question

If I have a dataset and find that sd=sqrt(mean), what's the criteria for favoring one distribution vs the other?



* * * 
