---
title: 'Week 12: MongoDB'
author: "Mauricio Alarcon"
date: "April 18, 2015"
output: html_document
---

##Summary

In this excercise we will be playing with the [NYC Public High School Program Data](https://data.cityofnewyork.us/Education/NYC-Public-High-School-Program-Data/mreg-rk5p) from the NYC OpenData Catalog.


The goal is to:

* Import the catalog into MongoDB
* Load the catalog into R
* Present data including the following attributes:
  * Programs by Borough
  * Programs by Interest Area 

## The Libraries

```{r, warning=FALSE, message=FALSE}
library(RCurl)
library(rjson)
library(knitr)
library(rmongodb)
library(plyr)
library(dplyr)
```

## Connecting to MongoDB

```{r}
mongo <- mongo.create()
coll <- "test.nyc_data_json"
```


## Importing the data into MongoDB

Unfortunatelly the mongoimport command has it's limitations. In the process of experimenting with the data I stumbled upon a couple of limitations:

* Unable to import data with greater than 16MB
* The schema of the data source needed some tidying.

Therefore, I decided to import the data into Mongo thru this R-script.

```{r}
# url to data
json_data_url <- "https://data.cityofnewyork.us/api/views/mreg-rk5p/rows.json?accessType=DOWNLOAD"
# get the RAW JSON data
json_data <- getURL(json_data_url)

# Parse into a list
nyc_school_Programs <- fromJSON(json_data)

# Get the column names form the dataset
nyc_data_columns <- nyc_school_Programs[["meta"]][["view"]][["columns"]]

colnames <- c()
for (i in nyc_data_columns){
  colnames <- c(colnames,i[["name"]])
}

# Get the data from the dataset
nyc_data_data <- nyc_school_Programs[["data"]]

#turn data into dataframe

nyc_data <- data.frame(matrix(unlist(nyc_data_data), ncol=length(colnames), byrow=T))

colnames(nyc_data) <- colnames

#importing the data into JSON
# convert data back via DF -> JSON -> BSON -> MONGODB


nyc_data_json<-apply(nyc_data,1,function(x) toJSON(x))

#delete all previously imported entries
mongo.remove(mongo, coll)

#export to mongo
apply(nyc_data,1,function(x) {mongo.insert(mongo
                                                    , coll
                                                    ,mongo.bson.from.JSON(
                                                      toJSON(x)
                                                      )
                                                    )
                                       }
      )

```


# Programs by Borough


```{r}

if(mongo.is.connected(mongo) == TRUE) {
  pipe_1 <- mongo.bson.from.JSON('
            {"$group":
              {"_id":"$Borough","Programs":{"$sum":1}}
            }')
  cmd_list <- list(pipe_1)
  totalPrograms <- mongo.aggregation(mongo, coll, cmd_list)
  
  ltotalPrograms <- mongo.bson.value(totalPrograms , "result")
  mtotalPrograms  <- sapply(ltotalPrograms , 
                    function(x) return( c(x$'_id',x$Programs)) )
  dtotalPrograms  <- as.data.frame(t(mtotalPrograms ))
  rownames(dtotalPrograms ) <- dtotalPrograms [,1]
  colnames(dtotalPrograms ) <- c("Borough", "Programs")
  head(dtotalPrograms)
  dtotalPrograms$Programs <- as.numeric(as.character(dtotalPrograms$Programs))

  plot(dtotalPrograms,main='High School Programs by Borough')

  kable(dtotalPrograms[c(2)])
}
```


# Programs by Interest Area


```{r}

if(mongo.is.connected(mongo) == TRUE) {
  pipe_1 <- mongo.bson.from.JSON('
            {"$group":
              {"_id":"$Interest Area","Programs":{"$sum":1}}
            }')
  cmd_list <- list(pipe_1)
  totalPrograms <- mongo.aggregation(mongo, coll, cmd_list)
  
  ltotalPrograms <- mongo.bson.value(totalPrograms , "result")
  mtotalPrograms  <- sapply(ltotalPrograms , 
                    function(x) return( c(x$'_id',x$Programs)) )
  dtotalPrograms  <- as.data.frame(t(mtotalPrograms ))
  rownames(dtotalPrograms ) <- dtotalPrograms [,1]
  colnames(dtotalPrograms ) <- c("Interest Area", "Programs")
  head(dtotalPrograms)
  dtotalPrograms$Programs <- as.numeric(as.character(dtotalPrograms$Programs))

  kable(dtotalPrograms[c(2)])
}
```

