---
title: 'Week 9: XML and JSON'
author: "Mauricio Alarcon"
date: "March 21, 2015"
output: html_document
---

##Summary

This is a warm up exercise to help you to get more familiar with the HTML, XML, and JSON file formats, and using packages to read these data formats for downstream use in R data frames. In the next two class weeks, we’ll be loading these file formats from the web, using web scraping and web APIs.

I selected three of my favorite books along with the following attributes:

* Title
* Authors
* Publish Date
* Hardcover Price
* Kindle Price

The data was saved in three different formats:

* HTML
* JSON
* XML

## The Libraries

```{r}
library(RCurl)
library(knitr)
library(rjson)
library(plyr)
library(XML)
```

## Getting JSON Data in R

```{r}
json_data_url <- "https://raw.githubusercontent.com/rmalarc/is607/master/assignment9/books.json"
json_data <- getURL(json_data_url)

books_json_data <- fromJSON(json_data)

books_json_df <- ldply(books_json_data,  function(x) { data.frame(x) } )

kable(books_json_df)

```

## Getting XML Data in R

```{r}
xml_data_url <- "https://raw.githubusercontent.com/rmalarc/is607/master/assignment9/books.xml"
xml_data <- getURL(xml_data_url)
books_xml_list <- xmlToList(xmlParse(xml_data))

books_xml_df <- ldply(books_xml_list, function(x) { data.frame(x) } )

kable(books_xml_df)
```


## Getting HTML Data in R

```{r}
html_data_url <- "https://raw.githubusercontent.com/rmalarc/is607/master/assignment9/books.html"
html_data <- getURL(html_data_url)

#get all the tables in the data
book_tables <- readHTMLTable(html_data)

# get the row number for all tables in html
n.rows <- unlist(lapply(book_tables, function(t) dim(t)[1]))

# select the table with the most rows
books_html_df<- book_tables[[which.max(n.rows)]]


kable(books_html_df)

```

##Conclusion

Even though there are R-libraries available for loading XML, HMTL, and JSON, the resulting R-dataframe schema varies slightly.

In addition, the different data-sources have unique challenges, particularly with HTML where handling documents that contain multiple tables can be a challenge.
